Symmetric Circuits for Rank Gates

### 2018-02-05 sections/translating-formulas-to-cicuits.tex
February 5 2018

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_q\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

@@

% Moreover, if we run merge on a circuit $C$ and a set $H$ then if $C$ has
% unique labels and $H$ has the property that there is no gate $g$ in $C$ such
% that two elements of $H$ are in $H_g$ then if $C$ has unique labels then so
% does $C'$. Moreover, if $C$ is symmetric then so is $C'$.

@@

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemize}
\end{definition}

@@

% Furthermore, $\op{make-injective}(C, h) = C'$ and there exists a symmetric
% gate $g \in C$ such that $\vert L(g)^{-1}(h) \vert > 1$ then the number of
% non-injective symmetric gates in $C'$ is strictly less than the number of
% non-injective symmetric gates in $C'$.

@@

% $ $g$ in $C$ such that $W(h,g)$, let $g_{h_1}$ be the first gate in the
% circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1) \vert \geq
% L(g')^{-1}(h_1)$

% We define $C'$ as follows. Let $F$ be a function that takes in an ordered
% circuit $C = \langle G, \Omega, \Sigma, \Lambda, L \rangle$ and outputs a
% circuit $C* := \langle G, \Omega, \Sigma, \Lambda, L \rangle$ defined as
% follows. If for all $g \in G$, $L(g)$ is an injection then $F(C) = C$ and if
% not then $F(C)$ is defined as follows.

% Let $j \in [\depth(C)]$, and let ${G_0}' = G$, ${\Sigma_0}' = \Sigma$ and
% ${L}_0' = L$.

% Let $S_j$ be the set of all gates of height $j$ that are symmetric-gates. For
% each $h_1 \in \bigcup_{g \in S_j} H_g$ let $g_{h_1} \in S_j$ be the first gate
% in the circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1)
% \vert \geq L(g')^{-1}(h_1)$. Let $(a_1, \ldots, a_p) := L(g_{h_1})^{-1}(h_1)$
% (ordered according the to order on the circuit). We add to $G'$ the gates
% $\{h_2, \ldots, h_p\}$ and for each $i \in \{2, \ldots, p\}$ we set
% $\Sigma'(h_i)$ such that $h_i$ is a $\land$ gate with a single output and
% $L'(h_i)$ such that $h_i$ takes only $h_{i-1}$ as an input. For each $g' \in
% S$ such that $h_1 \in H_g$ let $\{ b_1, \ldots, b_q \} := L(g)^{-1}(h_1)$. For
% each $i \in [q]$ let $L'(g')(b_i) = h_i$. Iterate $k$ and repeat this
% procedure until $k$ exceeds the depth of the circuit.

% \begin{definition}
%   The operation \emph{separate-labels} takes as input an ordered $(\mathbb{B},
%   \tau)$-circuit $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ and
%   outputs the circuit $C' = \langle G', \Omega', \Sigma', \Lambda', L'
%   \rangle$. We define $C'$ as follows. Let $j \in [\depth(C)]$, and let
%   ${G_0}' = G$, ${\Sigma_0}' = \Sigma$ and ${L}_0' = L$. Let $S_j$ be the set
%   of all gates of height $j$ that are symmetric-gates. For each $h_1 \in
%   \bigcup_{g \in S_j} H_g$ let $g_{h_1} \in S_j$ be the first gate in the
%   circuit such that for all $g' \in S_j$, $\vert L(g_{h_1})^{-1}(h_1) \vert
%   \geq L(g')^{-1}(h_1)$. Let $(a_1, \ldots, a_p) := L(g_{h_1})^{-1}(h_1)$
%   (ordered according the to order on the circuit). We add to $G'$ the gates
%   $\{h_2, \ldots, h_p\}$ and for each $i \in \{2, \ldots, p\}$ we set
%   $\Sigma'(h_i)$ such that $h_i$ is a $\land$ gate with a single output and
%   $L'(h_i)$ such that $h_i$ takes only $h_{i-1}$ as an input. For each $g' \in
%   S$ such that $h_1 \in H_g$ let $\{ b_1, \ldots, b_q \} := L(g)^{-1}(h_1)$.
%   For each $i \in [q]$ let $L'(g')(b_i) = h_i$. Iterate $k$ and repeat this
%   procedure until $k$ exceeds the depth of the circuit.
% \end{definition}

% Notice that for a given circuit $C$ and a symmetric-gate $g$ in $C$, running
% `merge' on each equivalence class of syntactically equivalent gates input to
% $g$ and then running `separate-labels' on the resultant circuit produces
% circuit $C'$ computing the same function as $C$ but the gate $g$ in $C$ has
% unique labels.

@@

% \begin{definition}
%   The algorithm \emph{tag} takes in a $(\mathbb{B}, \tau)$-circuit $C_n :=
%   \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a relation symbol $R \in
%   \tau$, a gate $g \in G$ and a tuple $\vec{a} \in [n]^k$ for some $k$ and
%   outputs the circuit $C_n' = \langle G', \Omega', \Sigma', \Lambda',
%   L'\rangle$ defined by the following algorithm. Let $G' = G \uplus \{\land_g,
%   \land_{g'}, \land_1, \ldots , \land_k\}$, $\Omega' = \Omega$, and $\Lambda'
%   = \Lambda$. Define $L'$ and $\Sigma'$ as follows. For each $g \in G$
%   $\Sigma'(g) = \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots ,
%   \land_k \}$, $\Sigma(g)$ is an $\AND$-gate that takes in two inputs.
%   $\Sigma(\land_g)$ is an $\AND$-gate that takes in a single input. Set $g$ to
%   be the lone input to $\land_g$. Set the two inputs to $\land_0$ to be
%   $\land_g$ and $\land_1$. For each $i \in [k-1]$ set the inputs of $\land_i$
%   to be $\land_{i+1}$ and the gate $g_i \in G'$ such that $\Sigma(g_i) = R$
%   and $\Lambda^{-1}(g_i) = \vec{a}_i$, where $\vec{a}_i$ is the $k$-tuple
%   $(\vec{a}(i), \ldots, \vec{a}(i))$. Set the inputs of $\land_k$ to be the
%   constant gate $1$ and the gate $g_k$ such that $\Sigma(g) = R$ and
%   $\Lambda^{-1}(g_k) = \vec{a}_k$.

%   For each gate $h \in G \setminus \{g\}$ and $\vec{b} \in \ind(h)$, if
%   $L(h)(\vec{b}) = g$ then let $L'(h)(\vec{b}) = \land_0$, otherwise let
%   $L'(h)(\vec{b}) = L(h)(\vec{b})$.

% \end{definition}

@@

  Suppose there is a linear order on $H$, and write $H = \{h_1, \ldots, h_t\}$.
  Let $C_0 := C$ and for all $i \in [t]$ let $C_i := \op{tag}(C_{i-1}, R, h_i,
  T(h_i))$. Then $C' = C_{t}$.

  % such that $G' = G \uplus \{\land_g, \land_{g'}, \land_1, \ldots ,
  % \land_k\}$,
  % $\Omega' = \Omega$, and $\Lambda' = \Lambda$. For each $g \in G$ $\Sigma'(g)
  % =
  % \Sigma(g)$. For each $g \in \{\land_0, \land_1, \ldots , \land_k \}$,
  % $\Sigma(g)$ is an $\AND$-gate that takes in two inputs. $\Sigma(\land_g)$ is
  % an $\AND$-gate that takes in a single input, $g$ is the lone input to
  % $\land_g$, and $\land_0$ has as input $\land_g$ and $\land_1$. For each $i
  % \in
  % [k-1]$ the inputs of $\land_i$ are $\land_{i+1}$ and the gate $g_i \in G'$
  % such that $\Sigma(g_i) = R$ and $\Lambda^{-1}(g_i) = \vec{a}_i$, where
  % $\vec{a}_i$ is the $k$-tuple $(\vec{a}(i), \ldots, \vec{a}(i))$. The inputs
  % of
  % $\land_k$ are the constant gate $1$ and the gate $g_k$ such that $\Sigma(g)
  % =
  % R$ and $\Lambda^{-1}(g_k) = \vec{a}_k$. For each gate $h \in G \setminus
  % \{g\}$ and $\vec{b} \in \ind(h)$, if $L(h)(\vec{b}) = g$ then
  % $L'(h)(\vec{b})
  % = \land_0$, otherwise $L'(h)(\vec{b}) = L(h)(\vec{b})$.
\end{definition}

@@

% \begin{definition}
%   The function $\op{concatenate}$ maps a $(\mathbb{B}, \tau)$-circuit $C :=
%   \langle G, \Omega, \Sigma, \Lambda, L \rangle$, a sequence of relation
%   sybmols $\tau_R := (R_1, \ldots, R_t)$ such that $\tau_R \subseteq \tau$, a
%   sequence of circuits $\mathcal{C} := ( C^1, \ldots, C^t)$, where for each $i
%   \in [t]$, $C^i = \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is
%   a $(\mathbb{B}_i, \tau_i)$-circuit, $\tau_i \not\subset \tau_R$ and $C^i$
%   computes a query of arity $\arty(R_i)$, and $R$ is a relation symbol in
%   $(\tau \setminus \tau_R) \cup (\bigcup_{1 \leq i \leq t},\tau_i))$ to a
%   circuit $\op{concatenate}(C, \mathcal{C}, \tau_R, R) := C' := \langle G',
%   \Omega', \Sigma', \Lambda', L' \rangle$. Note that all circuits take in
%   structures of size $n$. The circuit $C'$ is defined as follows.

%   Let ${C^0}' := \langle G^0, \Omega^0, \Sigma^0, \Lambda^0, L^0 \rangle :=
%   C$. For each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus
%   \tau_R$. For each $i \in [t]$ let ${C^{i}}' := \langle G_i', \Omega_i',
%   \Sigma_i', \Lambda_i', L_i' \rangle := \op{insert} (C^i, {C^{i-1}}')$ be
%   defined as follows. Let $D^{i-1}:= \op{insert} (C^i, {C^{i-1}}')$. Let
%   $G_{R_i}$ be the set of all relational gates in $D^{i-1}$ labelled by the
%   relational symbol in $R_i$. Let ${C^{i}}'$ be the circuit formed by
%   replacing each gate in $g \in G_{R_i}$ with the gate
%   $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ (by repeated calls to the
%   $op{replace}$ function) and using the $\op{tag}$ function on each gate
%   $\Omega_i (\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
%   $\Lambda_{i-1}'(g)$ and the relation symbol $V_i$. We then let $C' :=
%   {C^{t}}'$.

%   %   For each relational gate $g$ in $D^{i-1}$ labelled by the symbol
%   %   $R_i$,
%   %   $\op{replace}$ $g$ with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in
%   %   $D^{i-1}$
%   %   and $\op{tag}$ the gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$
%   %   using
%   %   the
%   %   tuple $\Lambda_{i-1}'(g)$ and the relation symbol $R$. Output the
%   %   circuit
%   %   $C'
%   %   := {C^t}'$.
% \end{definition}
% GW: do this tomorrow.

% \begin{definition}
%   The operation $\op{concatenate-circuits}(C, \mathcal{C}, \tau_R) = C'$,
%   where $C := \langle G, \Omega, \Sigma, \Lambda_0, L_0 \rangle$ is a
%   $(\mathbb{B}, \tau)$-circuit, $\mathcal{C}$ is a finite set of circuits
%   takes as input a set of relation symbols $\tau_R := \{R_1, \ldots, R_t\}$
%   and $t+1$ circuits $\{C^0, \ldots, C^{t}\}$, where $C^0 = \langle G_0,
%   \Omega_0, \Sigma_0, \Lambda_0, L_0 \rangle$ is a $(\mathbb{B},
%   \tau)$-circuit with $\tau_R \subseteq \tau$ and for each $i \in [t]$, $C^i =
%   \langle G_i, \Omega_i, \Sigma_i, \Lambda_i, L_i \rangle$ is a
%   $(\mathbb{B}_i, \tau_i)$-circuit computing a query of arity $\arty(R_i)$.
%   The algorithm outputs, for a given input, a circuit $C = \langle G, \Omega,
%   \Sigma, \Lambda, L\rangle$ constructed as follows. Let ${C^0}' := C^0$. For
%   each $i \in [t]$ fix a relation symbol $V_i \in \tau_i \setminus \tau_R$.
%   For each $i \in [t]$ construct the circuit ${C^{i}}' = \langle G_i',
%   \Omega_i', \Sigma_i', \Lambda_i', L_i' \rangle $ by `inserting' $C^i$ into
%   ${C^{i-1}}'$, and call the resultant circuit ${D^{i-1}}'$. For each
%   relational gate $g$ in $D^{i-1}$ labelled by the symbol $R_i$, `replace' $g$
%   with the gate $\Omega_{i}(\Lambda_{i-1}'(g))$ in $D^{i-1}$ and `tag' the
%   gate $\Omega_i(\Lambda_{i-1}'(g))$ in $D^{i-1}$ using the tuple
%   $\Lambda_{i-1}'(g)$ and the relation symbol $V_i$. Output the circuit $C :=
%   {C^t}'$.
% \end{definition}

@@

It is easy to see that $\op{compose-and-tag}$ can be computed in time polynomial
in the sizes of the input circuits. Furthermore, if the input sequences of
circuits and relations are both singletons, then if the input circuits are all
transparent (resp. have unique labels) then the output circuit is transparent
(resp. have unique labels).

@@

% From Lemma \cite{} we may assume, without a loss of generality, that $(C_n)_{n
% \in nats}$ is a family of rigid circuits (and hence has unique labels).

@@

\begin{proof}
  We say a formula $\psi(\vec{x})$ is \emph{circuit-translatable} if there is a
  $P$-uniform family of transparent symmetric rank-circuits computing the query
  as the one defined by $\psi(\vec{x})$. If a formula $\psi(\vec{x})$ is
  circuit-translatable we let $(C^\psi_n)_{n \in \nats}$ denote a $P$-uniform
  family of transparent symmetric rank-circuits computing the same query as
  defined by $\psi(\vec{x})$.

  Let $\theta(\vec{x}) \in \FPR[\tau]$. We prove that $\theta(\vec{x})$ is
  circuit-translatable by structural induction. That is, we prove that if every
  sub-formula of $\theta(\vec{x})$ is circuit-translatable then
  $\theta(\vec{x})$ is circuit-translatable. Since it is obvious that every
  atomic formula in $\FPR[\tau]$ is circuit-translatable, the result thus
  follows. We already have that the formulas of $\FPC[\tau]$ are
  circuit-translatable (Anderson and Dawar~\cite{AndersonD17}), and in fact we
  may assume, without a loss of generality, that this translation yields a
  family of transparent circuits with unique labels (and unique extensions).

  We begin by showing that for any $\theta(\vec{x}) \in \FPR[\tau]$, there
  exists a $P$-uniform family of formulas $(\theta(\vec{x}))_{n \in \nats}$ such
  that for each $n \in \nats$, $\theta_n(\vec{x})$ is a first order formula with
  counting and rank quantifiers and is equivalent to $\theta(\vec{x})$ on
  structures of size $n$. This fact follows immediately from three observations.
  First, we note that for any $\phi \in FPR[\tau]$ of the form $\lfp{V,
    \vec{y}}{\psi(\vec{x}, \vec{y}, V)}(\vec{x})$ , there is a mapping $n
  \mapsto \phi_n$, computable in time polynomial in $n$, such that for all $n
  \in \nats$, both $\phi_n$ and $\phi$ are equivalent on structures of size $n$,
  and $\phi_n$ contains no least-fixed-point operator. The formula $\phi_n$ is
  define from $\phi$ by `unfurling' the least-fixed-point operator, producing,
  for a given $n \in \nats$, a formula that explicitly implements the definition
  of the operator in the obvious manner (for details on this implementation see
  \ref{}). This construction of $\phi_n$ involves adding a constant number of
  terms to $\phi$ for each iteration involved in computing the
  least-fixed-point. Since there are at most $n^k$ such iterations, where $k$ is
  the width of the formula $\psi$, the construction of $\phi_n$ can be completed
  in time polynomial in $n$.

  Similar arguments holds for number terms and rank operators. We have
  from~\cite{} that for any formula $\phi$ containing number-terms there is a
  $P$-uniform family of formulas $(\phi_n)_{n \in \nats}$ containing counting
  quantifiers and no number-terms, and such that for all $n \in \nats$, $\phi$
  and $\phi_n$ are equivalent on structures of size $n$. We have from Dawar et
  al.~\cite{Dawar09logicswith} that for every formula $\phi \in \FPR[\tau]$
  containing rank operators, there exists a $P$-uniform family that takes as
  input a natural number $n$ and outputs a formula $\phi_n$ with rank
  quantifiers and no rank operators, and such that for all $n \in \nats$, $\phi$
  and $\phi_n$ are equivalent on structures of size $n$.

  The algorithm given by executing each of the above three algorithms in
  succession computes the required mapping $n \mapsto \theta_n(\vec{x})$ in time
  polynomial in $n$, thus giving us the required $P$-uniform family
  $(\theta_n(\vec{x}))_{n \in \nats}$.

  It follows that the number of sub-formulas of $\theta_n$

  It follows that in order to prove that $\theta(\vec{x})$ is
  circuit-translatable it is sufficient to prove that there exists a function
  that maps $\theta_n$ (for any $n \in \nats$) to $C^{\theta}_n$, a transparent
  symmetric rank-circuit that computes the same query as defined by
  $\theta_n(\vec{x})$ (and so $\theta(\vec{x})$) for structures of size $n$, and
  is computable in time polynomial in in both $n$ and the number of sub-formulas
  in $\theta_n$. This is sufficient as the number of sub-formulas in $\theta_n$
  is bounded by some polynomial in $n$, and so composing functions, we can
  compute the map that takes $n$ to $\theta_n$ to $C^{\theta}_n$ in time
  polynomial in $n$.

  We define an algorithm for constructing $C^{\theta}_n$ given
  $\theta_n(\vec{x})$ by recursion on the structure of the input formula. The
  first note that if $\theta_n(\vec{x})$ an atomic formula then there is a
  trivial construction of $C^{\theta}_n$.

  We consider the conjunctive case. Suppose $\phi_1, \phi_2 \in \FPR[\tau]$ are
  circuit-translatable and $\theta (\vec{x}) = \phi_1(\vec{x}) \land
  \phi_2(\vec{x})$. Let $R_{\phi_1}$ and $R_{\phi_2}$ be two $\vert \vec{x}
  \vert$-ary relation symbols not in $\tau$. Let $\theta' (\vec{x}) :=
  R_{\phi_1}(\vec{x}) \land R_{\phi_2}(\vec{x})$. Let $n \in \nats$. Since
  $\theta'(\vec{x})$ is a formula of $\FPC[\tau]$, we have from Anderson and
  Dawar~\cite{AndersonD17} that $\theta'(\vec{x})$ is circuit-translatable. We
  let $C^{\theta}_n := \op{compose}(C^{\theta'}, (C^{\phi_1}, C^{\phi_2}),
  (R_{\phi_1}, R_{\phi_2}))$.

  Since all internal gates in $C^{\theta'}$ are symmetric, and both $C^{\phi_1}$
  and $C^{\phi_2}$ are transparent, it follows that $C^{\theta}_n$ is
  transparent. Moreover, it is easy to see that the construction of
  $C^{\theta}_n$ can be carried implemented so as to run in time polynomial in
  $n$, and that $(C^{\theta}_n)_{n \in \nats}$ computes the same query defined
  by $\theta(\vec{x})$. It follows that $\theta(\vec{x})$ is
  circuit-translatable.

  % Notice that $D^{\theta}_n$ computes the same relation as defined by
  % $\theta(\vec{x})$ for structures of size $n$. Moreover, since the remaining
  % two operations preserve the function computed, it follows $C^{\theta}_n$
  % computes the same relation as defined by $\theta(\vec{x})$ for structures of
  % size $n$. Furthermore, since $C^{\phi_1}_n$, $C^{\phi_2}_n$ and
  % $C^{\theta'}_n$ are symmetric so is $C^{\theta}_n$. Lastly, notice that all
  % non-output gates in $D^{\theta}_n$ have unique labels. After the
  % $\op{merge-all}$ and $\op{make-injective-all}$ operation the output gates
  % have
  % unique labels, and thus $C^{\theta}_n$ has unique labels. This construction
  % takes in a natural number $n$ and produces the required circuit
  % $C^{\theta}_n$. Moreover, it is easy to see that this construction can be
  % implement as an algorithm running in time polynomial in $n$. It follows that
  % $\theta(\vec{x})$ is circuit-translatable.

  The approach used for resolving the conjunctive case can be thought of as
  having three steps. First, replacing each sub-formula in $\theta$ with a
  relation symbol, resulting in a formula $\theta'$ of $\FPC$. Second, using the
  algorithms from the inductive hypothesis' (and the results of Anderson and
  Dawar) on $\theta'$ and the sub-formulas, producing appropriate families for
  $\theta'$ and each sub-formula of $\theta$. Third, for each $n \in \nats$
  composing the circuit $C^{\theta'}_n$ with the circuits for each sub-formula,
  producing the the circuit $C^{\theta}_n$. Since this process can be executed
  in time polynomial in $n$, we conclude that $\theta$ is circuit-translatable.

  This same approach can be used to handle many other cases. For
  circuit-translatable $\phi, \phi_1, \phi_2 \in \FPR[\tau]$, and $k \in \nats$,
  similar constructions can be used to prove that if $\theta(\vec{x}) \in
  \FPR[\tau]$ is of the form $\phi_1(\vec{x}) \lor \phi_2(\vec{x})$, $\neg
  \phi(\vec{x})$, $\forall y. \phi (\vec{x}, y)$, $\exists y. \phi (\vec{x}, y)$
  or $C^k y. \phi(\vec{x}, y)$ then $\theta(\vec{x})$ is circuit-translatable.

>>>>

\begin{definition}
  Let $C := \langle G, \Omega, \Sigma , \Lambda, L \rangle$ be a circuit and $H
  := \{h_1, \ldots, h_p\}$ be a set of syntactically equivalent gates. We let
  $\op{merge}(C, H) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, where
  \begin{itemize}
    \setlength\itemsep{0mm}
  \item $G' = G \setminus \{h_2, \ldots, h_p\}$,
  \item $\Omega' = \Omega$,
  \item $\Sigma' = \restr{\Sigma}{G'}$,
  \item $\Lambda' = \Lambda$, and
  \item For all $g \in G'$ and $\vec{a} \in Ind(g)$ if $L(g) (\vec{a}) \in H$
    then $L'(g)(\vec{a}) = h_1$ and if $L(g)(\vec{a}) \notin H$ then
    $L'(g)(\vec{a}) = L(g)(\vec{a})$.
  \end{itemize}
\end{definition}

@@

@@

\begin{definition}
  \label{def:make-injective}
  The function $\op{make-injective}$ maps a circuit $C := \langle G, \Omega,
  \Sigma , \Lambda, L \rangle$ and a gate $h \in G$ to the circuit
  $\op{make-injective}(C, h) := C' := \langle G', \Omega', \Sigma', \Lambda', L'
  \rangle$, defined as follows. Let $c = \max \{\vert L(g)^{-1}(h)\vert : g \in
  W(h, \cdot)\}$. If $c = 1$ then $C' := C$. If $c > 1$ then $C'$ is defined
  such that
  \begin{itemizens}
  \item $G' = G \uplus \{h_2, \ldots, h_{c-1}\}$,
  \item $\Omega' = \Omega$,
  \item For each $g \in G$ let $\Sigma' (g) = \Sigma(g)$, otherwise let
    $\Sigma'(g)$ be a single-input $\AND$-gate.
  \item $\Lambda'= \Lambda$
  \item For each $g \in G'$ if $g \in G \setminus W(h, \cdot)$ then $L'(g) =
    L(g)$. If $g \in W(h, \cdot)$ let $\{\vec{a}_1, \ldots, \vec{a}_k\} :=
    L(g)^{-1}(h)$ and for each $\vec{a} \in \ind(g)$ if $L(g)(\vec{a}) \neq h$
    then $L'(g)(\vec{a}) = L(g)(\vec{a})$ and if $\vec{a} = \vec{a_i} \in
    L(g)^{-1}(h)$ then let $L'(g)(\vec{a}) = h_i$.
  \end{itemizens}
\end{definition}

@@

@@

@@

@@

  Suppose there is a linear order on $H$, and write $H = \{h_1, \ldots, h_t\}$.
  Let $C_0 := C$ and for all $i \in [t]$ let $C_i := \op{tag}(C_{i-1}, R, h_i,
  T(h_i))$. Then $C' = C_{t}$.
\end{definition}

@@

@@

It is easy to see that $\op{compose-and-tag}$ can be computed in time polynomial
in the sun of the sizes of the input circuits. We also have that if the input
circuits are symmetric then the output circuits are symmetric. Furthermore, if
the input sequences of circuits and relations are both singletons, then if the
input circuits are all transparent (resp. have unique labels) then the output
circuit is transparent (resp. have unique labels).

@@

@@

\begin{proof}
  We say that a circuit $C$ \emph{translates} a formula $\psi(\vec{x})$ for $n
  \in \nats$ if $C$ is a transparent symmetric rank-circuit of order $n$ and $C$
  computes the query defined by $\psi(\vec{x})$ on structures of size $n$. We
  say a family of circuits $(C_n)_{n \in \nats}$ \emph{translates} a formula
  $\psi (\vec{x})$ if $(C_n)_{n \in \nats}$ is $\PT$-uniform and for all $n \in
  \nats$, $C_n$ translates $\psi(\vec{x})$ for $n$.

  We have from Anderson and Dawar~\cite{AndersonD17} that for every formula
  $\phi(\vec{x}) \in \FPC[\tau]$ there is a $\PT$-uniform family of symmetric
  circuits $(C^{\phi}_n)_{n \in \nats}$ with unique labels and symmetric gates
  that translates $\phi(\vec{x})$.

  Let $\theta(\vec{x}) \in \FPR[\tau]$ and $t := \width (\theta)$. We begin by
  showing that there exists a $P$-uniform family of formulas
  $(\theta(\vec{x}))_{n \in \nats}$ such that for each $n \in \nats$, we have
  $\theta_n(\vec{x}) \in \FOrk[\tau]$, $\width(\theta_n) \leq 4 t$, and
  $\theta_n(\vec{x})$ is equivalent to $\theta(\vec{x})$ on structures of size
  $n$. This result follows from two observations.

  First, using the translation in~\cite{Dawar09logicswith}, we have that for any
  $\phi(\vec{x}) \in \FPR[\tau]$ of the form $\lfp{V, \vec{y}}{\psi(\vec{x},
    \vec{y}, V)}(\vec{x})$ , there is a mapping $n \mapsto \phi_n$, computable
  in time polynomial in both $n$ and the size of $\phi(\vec{x})$ such that for
  all $n \in \nats$, we have $\phi_n(\vec{x}) \in \FOR[\tau]$, $\width (\phi_n)
  \leq 2 \cdot \width (\phi)$, and $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are
  equivalent on structures of size $n$.

  Second, using the translation in~\cite{libkin2004elements}, we have that if
  $\phi (\vec{x}) \in \FOR[\tau]$ is of the form $\rank (\vec{y}\vec{\mu} \leq
  \vec{t}, \vec{z}\vec{\nu} \leq \vec{s}, \pi \leq r) . \psi (\vec{y}\vec{\mu},
  \vec{z} \vec{\nu})$, then there is a mapping $n \mapsto \phi_n$ computable in
  time polynomial in both $n$ and the size of $\phi(\vec{x})$ such that for all
  $n \in \nats$, $\phi_n(\vec{x}) \in \FOrk[\tau]$, $\width(\phi_n) \leq 2 \cdot
  \width(\phi)$, and $\phi_n(\vec{x})$ and $\phi(\vec{x})$ are equivalent on
  structures of size $n$.

  We thus derive the required $\PT$-uniform family $(\theta_n(\vec{x}))_{n \in
    \nats}$ by applying the first translation and then the second.

  Fix $n \in \nats$. We recursively define a function $T_n$ that maps $\psi
  (\vec{y})$, a sub-formula of $\theta_n(\vec{x})$, to a circuit $C^{\psi}_n$,
  where $C^{\psi}_n$ translates $\psi(\vec{y})$ at $n$. We then have that for
  all $n \in \nats$, $T_n(\theta_n(\vec{x}))$ translates $\theta_n$ (and hence
  $\theta$) at $n$.

  We first consider the base case. Suppose $\psi(\vec{y})$ is an atomic formula.
  Then $\psi(\vec{y}) \in \FPC[\tau]$, and so, using the translation defined by
  Anderson and Dawar~\cite{AndersonD17}, there exists a family of circuits
  $(C^{\psi}_n)_{n \in \nats}$ that translates $\psi(\vec{y})$. Let
  $T_n(\psi(\vec{y})) := C^{\psi}_n$.

  We now consider the conjunctive case. Suppose $\psi(\vec{y})$ is a sub-formula
  of $\theta_n(\vec{x})$ such that $\psi (\vec{y}) = \phi_1(\vec{y}) \land
  \phi_2(\vec{y})$. Let $R_{\phi_1}$ and $R_{\phi_2}$ be two $\vert \vec{y}
  \vert$-ary relation symbols not in $\tau$ and let $\psi' (\vec{y}) :=
  R_{\phi_1}(\vec{y}) \land R_{\phi_2}(\vec{y})$. Then $\psi'(\vec{y}) \in
  \FPC[\tau]$ and so, using the translation defined by Anderson and Dawar, there
  exists a circuit family $(C^{\psi'}_n)_{n \in \nats}$ that translates
  $\psi'(\vec{y})$. Let $T_n(\psi (\vec{y})) := \op{compose}(C^{\psi'}_n, (
  T_n(\phi_1(\vec{y})), T_n(\phi_2(\vec{y}))), (R_{\phi_1}, R_{\phi_2}))$. We
  note that $T_n(\psi(\vec{y}))$ translates $\psi (\vec{y})$ at $n$ if
  $T_n(\phi_1 (\vec{y}))$ and $T_n(\phi_2(\vec{y}))$ translate $\phi_1(\vec{y})$
  and $\phi_2(\vec{y})$ at $n$, respectively.

  % We also note that this construction gives us that, if we exclude the cost of
  % the recrive calls, then $n \mapsto T_n(\psi(\vec{y}))$ is computable in time
  % polynomial in $n$. Notice, though, that this polynomial does depend on
  % $\psi'(\vec{y})$, i.e. the form of $\psi (\vec{y})$.

  The approach used in the conjunctive case can be summarised as follows. Given
  a sub-formula $\psi (\vec{y})$, we construct an auxiliary formula
  $\psi'(\vec{y})$ in which all the sub-formulas of $\psi(\vec{y})$ are replaced
  with relation symbols. We then construct a circuit $C^{\psi'}$ that translates
  $\psi'(\vec{y})$ at $n$ using the translation defined by Anderson and Dawar.
  Finally, we recursively construct circuits that translate the sub-formulas of
  $\psi(\vec{y})$, and compose these circuits with $C^{\psi'}$.

  This approach can be used in order to handle every case with the exception of
  rank. In other words, we can use a similar approach to define
  $T_n(\psi(\vec{y}))$ where $\psi(\vec{y})$ is a sub-formula of
  $\theta_n(\vec{x})$ of the form $\phi_1(\vec{y}) \lor \phi_2(\vec{y})$, $\neg
  \phi(\vec{y})$, $\forall z. \phi (\vec{y}, z)$, or $\exists z. \phi (\vec{y},
  z)$.

  We now handle the rank case. The approach is quite similar, except in this
  case we cannot use the translation defined by Anderson and Dawar in order to
  translate the auxiliary formula, and so must define this translation
  explicitly. Moreover, rather than just composing the circuits, we also add a
  small gadget in order to ensure the resultant circuit is transparent.
